\documentclass[../main/SI3_manual]{subfiles}

\begin{document}

% ================================================================
% Dummy Chapter - To do list
% ================================================================

\chapter*{To-do-list for documentation}
\label{chap:TDL}

\begin{itemize}

\item \textcolor{gray}{Documentation infrastructure ready by Jul 19, 2018.}

\item \textcolor{gray}{Abstract: v1 written.}

\item \textcolor{gray}{Disclaimer: v1 written.}

\item Introduction: list of chapters and change between realeases to be written (1h)

\item \textcolor{gray}{1. Model basics: v1 written.}

\item \textcolor{gray}{Add namelist infrastructure as in NEMO doc (done)}

\item \textcolor{gray}{2. Time, space and thickness space domain}

\item 3. Ice dynamics: to be written (Clem \& Martin, 3h)

\item 4. Ice transport: UM5 description to be written (Clem, 3h)

\item 5. Ridging and rafting: V1 from previous doc. Should be refreshed. A bit too long. All namelist parameters are not explained (namdyn\_rdgrft).

\item \textcolor{gray}{6. Radiative transfer: Full rewriting required (Martin \& Clem, 3-6h). Available elements from previous documentation, but mostly obsolete.}

\item 7. Thermodynamics: Lots of rewriting required. Available elements from previous documentation, but mostly obsolete (Martin, 6h)

\item 8. Interfaces: Full writing required 

\item 9. Output and diagnostics: Full writing required

\item 10. Single category (Martin, 3h). Full writing required.

\item 11. BDY-AGRIF (Clem, 3h). Full writing required.

\item 12. Miscellaneous. Full writing required.

\item Melt ponds are not currently considered in the plan.

\item Make sure all namelist parameters are properly described.

\item Peer-review by SIWG members.

\end{itemize}

\end{document}